---
title: "Social Capital and Crime <br>in Germany"
author: "Daniel Salgado Moreno and Lars Mehwald"
date: "4 December 2015"
output:
  slidy_presentation: default
  ioslides_presentation:
    logo: Supplementary/LogoHertie.png
---

## Outline
- Research Question
- Data
- Models
- Results

## Research Question
How do social networks and bonds in a given community determine the level of violent crime?

H0: *The more social capital in anygiven community, the less violent crime there would be observed.* 

H1: *The higher the foundations density is in a district, the lower the murder rate.* 

H2: *The lower the migration flow is in a district, the lower the murder rate.*  

H3: *The higher the voter turnout is in a district, the lower the murder rate.*

## Data

We gathered data from three main sources: 

1) German Federal Police's crime statistic ("Polizeiliche Kriminalstatistik" [PKS])  

2) The Regional Database Germany published by the Federal Statistical Office and the statistical Offices of the Lander

3) German Federal Foundations Association ("Bundesverband Deutscher Stiftungen")  

## Summary Statistics
```{r SummaryStatistics, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# Setting the working directory
library("repmis")
library("stargazer")
possible_dir <- c('D:/Eigene Dokumente/!1 Vorlesungen/!! WS 2015/Introduction to Collaborative Social Science Data Analysis/Assignment3', 
                  '~/HSoG/DataAnalysis/GitHub/Assignment3')
set_valid_wd(possible_dir)
rm(possible_dir)
# Loading data set from csv file
DistrictData <- read.csv(file="Analysis/data/DistrictData2013.csv")
# Removing ranking column (it was added in the saving process in DataMerging.R)
DistrictData <- DistrictData[,-1]
# Removing some variables
DistrictDataReduced <- DistrictData[,c(55,57,47,54,53,40,22,50,51,52)]
# Rounding

# Renaming variables
names(DistrictDataReduced)[names(DistrictDataReduced) == 'UnemployedPercentage'] <- 'UnemployedPerc'
names(DistrictDataReduced)[names(DistrictDataReduced) == 'FoundationsDensity100k'] <- 'FoundationsDens'
names(DistrictDataReduced)[names(DistrictDataReduced) == 'TurnoutPercentage'] <- 'TurnoutPerc'
# Creation of summary statistics
stargazer(DistrictDataReduced,
          nobs = FALSE,
          header = FALSE,
          digits=1,
          type = "html")
```

## Highest Murder Rates in Germany Districts
```{r Top10Murder, echo=FALSE, results='asis',warning=FALSE, message=FALSE}
source("Analysis/Graphs.R")

#top10MurderTab <- xtable(top10Murder, align = c("c","c","c","c"))
#print(top10MurderTab,type="html")

top10MurderMatrix <- as.matrix(top10Murder)
stargazer(top10MurderMatrix, header = FALSE, type="html", digits = 2)
```

## Correlation Matrix
```{r CorrChart, echo=FALSE, results='asis',warning=FALSE, message=FALSE}
source("Analysis/Graphs.R")
chart.Correlation(correlation.matrix, historgram=T)
```

## Murder Rate distribution
```{r Histogram Murder, echo=FALSE, results='asis',warning=FALSE, message=FALSE}
source("Analysis/Graphs.R")
plot(histMurderRate) 
```

## Comparing OLS models (controls suppressed)
```{r plotOLSMurder, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
source("Analysis/DataAnalysis.R")

stargazer(OLSMurderFoundations, OLSMurderFlow, OLSMurderTurnout, 
          type = "html",
          keep = c("FoundationsDensity100k", "FlowRate", "TurnoutPercentage"),
          header = FALSE, # important not to have stargazer information in markdown file 
          #title = "OLS regression models compared (controls suppressed)",
          digits = 2,
#          no.space = TRUE, # single.row = TRUE
          omit.stat = c("f", "ser"))
```
The control variables are: ForeignerRate, MarriageRate, MaleRate, YouthRate, UnemployedRate

## Poisson-based regression model (controls suppressed)
```{r plotPoissonMurder, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

source("Analysis/DataAnalysis.R")

stargazer(poisson.glm4,
          type = "html",
          keep = c("FoundationsDensity100k", "FlowRate", "TurnoutPercentage"),
          header = FALSE, # important not to have stargazer information in markdown file 
         # title = "Poisson-based regression model (controls suppressed)",
          digits = 2,
#          no.space = TRUE, # single.row = TRUE
          omit.stat = c("f", "ser"))
```
The control variables are: ForeignerRate, MarriageRate, MaleRate, YouthRate, UnemployedRate

## Negative Binomial (controls suppressed)
```{r plotNegbinMurder, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

source("Analysis/DataAnalysis.R")

stargazer(nb.glm4, 
          type = "html",
          keep = c("FoundationsDensity100k", "FlowRate", "TurnoutPercentage"),
          header = FALSE, # important not to have stargazer information in markdown file 
#          title = "Negative binominal regression model (controls suppressed)",
          digits = 2,
#          no.space = TRUE, # single.row = TRUE
          omit.stat = c("f", "ser"))
```
The control variables are: ForeignerRate, MarriageRate, MaleRate, YouthRate, UnemployedRate

## Incident Rates 
```{r NegbinSignificance, error=FALSE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("repmis")
library("MASS")
library("stargazer")

# Setting the commonly used working directory
possible_dir <- c('D:/Eigene Dokumente/!1 Vorlesungen/!! WS 2015/Introduction to Collaborative Social Science Data Analysis/Assignment3', 
                  '~/HSoG/DataAnalysis/GitHub/Assignment3')
set_valid_wd(possible_dir)
rm(possible_dir)

# Loading data set from csv file
DistrictData <- read.csv(file="Analysis/data/DistrictData2013.csv")

# Removing ranking column (it was added in the saving process in DataMerging.R)
DistrictData <- DistrictData[,-1]

####################################
# Data setup
# Declareing integre data for analysis
####################################

# Declaring distric Id as factor variablespo
DistrictData$district <- as.factor(DistrictData$district)
# Declaring all relevant variables for model integer
DistrictData$CrimeRate <- as.integer(DistrictData$CrimeRate)
DistrictData$FoundationsDensity100k <- as.integer(DistrictData$FoundationsDensity100k)
DistrictData$FlowRate <- as.integer(DistrictData$FlowRate)
DistrictData$TurnoutPercentage <- as.integer(DistrictData$TurnoutPercentage)
DistrictData$YouthRate <- as.integer(DistrictData$YouthRate)
DistrictData$MaleRate <- as.integer(DistrictData$MaleRate)
DistrictData$UnemployedPercentage <- as.integer(DistrictData$UnemployedPercentage)
DistrictData$MarriageRate <- as.integer(DistrictData$MarriageRate)
DistrictData$MurderRate <- as.integer(DistrictData$MurderRate)
DistrictData$ForeignerRate <- as.integer(DistrictData$ForeignerRate)

DistrictData$TotalPopulation <- as.integer(DistrictData$TotalPopulation)
DistrictData$murderAndManslaughter <- as.integer(DistrictData$murderAndManslaughter)
DistrictData$FoundationsTotal <- as.integer(DistrictData$FoundationsTotal)
DistrictData$OutflowTotal <- as.integer(DistrictData$OutflowTotal)
DistrictData$TurnoutPercentage <- as.integer(DistrictData$TurnoutPercentage)
DistrictData$ForeignersTotal <- as.integer(DistrictData$ForeignersTotal)
DistrictData$MalePopulation <- as.integer(DistrictData$MalePopulation)
DistrictData$Pop0to17 <- as.integer(DistrictData$Pop0to17)
DistrictData$Pop18to24 <- as.integer(DistrictData$Pop18to24)
DistrictData$Pop25to44 <- as.integer(DistrictData$Pop25to44)
DistrictData$PopOver65 <- as.integer(DistrictData$PopOver65)

# New name to Murder
DistrictData$Murder <- DistrictData$murderAndManslaughter

# New name to Couples: 
DistrictData$Couples <- DistrictData$HusbandAndWifeTotal

# FlowTotal
DistrictData$FlowTotal <- (DistrictData$OutflowTotal + DistrictData$InfluxTotal)

# Creating categorical variables from contoinous independent variables 
# Selection of cathegories based on quantiles
DistrictData$Foundations_cat <- cut(DistrictData$FoundationsTotal,
                                    breaks= c(0,15,26,44,1194), 
                                    labels= c('1stQu', '2ndQu', '3rdQu', '4thQu'))

nb.glm1 <- glm.nb(MurderRate ~ 
               FoundationsDensity100k +
               FlowRate +
               TurnoutPercentage +
               ForeignerRate +
               MarriageRate +
               MaleRate +
               YouthRate +
               UnemployedPercentage,
             DistrictData)

# Adding coefficients and confident intervals into new data frame 
est1 <- cbind(Estimate = coef(nb.glm1), 
              confint(nb.glm1, level=0.90),
              confint(nb.glm1, level=0.95),
              confint(nb.glm1, level=0.99))
est1 <- data.frame(est1)
est1 <- cbind(est1, 
              ifelse(sign(est1[2]) == sign(est1[3]), 1, 0),
              ifelse(sign(est1[4]) == sign(est1[5]), 1, 0),
              ifelse(sign(est1[6]) == sign(est1[7]), 1, 0))
est1 <- round(est1, 4)
est1 <- cbind(est1, 
              ifelse(est1[10] == 1, 3, 
              ifelse(est1[9] == 1, 2,
              ifelse(est1[8] == 1, 1, 0))))
est1 <- est1[,-c(2:10)]
est1 <- cbind(exp(est1[1]), est1[2])
names(est1) <- c("Coefficient", "NumberStars")
est1 <- round(est1, 4)
est1$NumberStars <- as.factor(est1$NumberStars)
est1$Stars <- ifelse(est1$NumberStars == 3, "+++", # &#9733; is HTML star
                            ifelse(est1$NumberStars == 2, "++",
                                   ifelse(est1$NumberStars == 1, "+", "")))
est1 <- est1[c(1,3)]
names(est1) <- c("IncidentRate", "_")
stargazer(est1, header = FALSE, summary= FALSE, type="html", digits = 4)
```

## <!-- last slide-->
<br><br><br><br><br><br>
<center><b>Thank you for your attention!<br>
Any questions?</b></center>

## Appendix 
### MC Simulations using 1st and 3rd Qu. as lower and upper bounds 
```{r MC Sim NB, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
source("Analysis/DataAnalysis.R")
plot(snb.out)
```

